SKETCH RECOGNITION AND UNDERSTANDING

### 2009-10-13 trunk/draft/ConFut.tex
FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL

%As you can see form the experiments PSO have proved a spuriously than other system. 
%\section{Future Work}
The next step in this research is to complete the clustering algorithm for fully automated sketch recognition. This will enable testing the sytsem for more than single symbol at a time. The segmentation algorithm may be more powerful if it can segments the stroke into more types of primitives other than the line and curve. Other area of modification can be the features extraction and classifier. Adding more spatial and structure features will improve classifications. After segmenting the stroke the system can generate a structural graph from segments and there spatial relations. This graph will be matched with the set of known symbol template to make the identification. Those results along with the SVM classifier results will enhance the system as a whole and achieve better recognition rate.  %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.  
\appendix
\chapter{Appendix: Step by Step Example Details}
%\addcontentsline{toc}{chapter}{Appendix: Step by Step Example Details}
\chaptermark{Appendix}
\markboth{Appendix}{Appendix}
\label{ChapterstepExample}%\chapter{Step by Step Example}
%In this chapter a step by step example is presented to clarify the presented system. Let us make an assumption that the user will draw a clock using two strokes. Figure \ref{fig:clock} show the two strokes and how the user draws them. The next steps demonstrate how output of the system after each step. The first section \ref{sec:stepseg} describe how the segmentation is done step by step. Section \ref{sec:steprec} shows the details of feature calculation and the classification process.

@@

%\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing} and Algorithm \ref{extractpdp}
\begin{algorithm}
\caption{Extracting $P_dp$}
\label{extractpdp}
\begin{verbatim}
  1. For each point in Stroke do the following:
  		1.a Compute distance from previous point.
  		1.b Compute Time difference from previous point and add it to Time_list.  
  		1.c Compute Velocity and add it to Vel_list. 
  		1.d Compute Direction and add it to Dir_list.
  	 	1.e Compute Curvature and add it to Cur_list.(Eq.4.1) 
  	end (for loop)
  2. For each  list computed in previous step do the follwing:   
  	 2.a Compute average of all points. 
  	 2.b set average as threshold.  
  	 2.c Create region from continuous points larger than threshold. 
  	 2.d for each region do the following. 
  	 	   2.d.1 find the maximum 
  	 	   2.d.1 Add point to P_dp 
  	  end
     end 
\end{verbatim}	
\end{algorithm}
%\item Ellipse Fitting: which is an attempt to fit as an ellipse as in Section \ref{sec:EllipseDetection} 
%he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
\begin{algorithm}
\caption{Ellipse Fitting Algorithm}
\label{ellipseFitAlg}
\begin{verbatim}
1. Initialize the axes of ellipse  and center by bounding box information
2. Repeat the following for few loops: 
	2.a Compute error in Equation  4.2 
	2.b Update center of ellipse and axis from error. 
3. Compute percent in Equation 4.4 
4. Computer eff in Equation 4.3 
5. if eff < 0.2  then fit as ellipse otherwise move to segment. 
\end{verbatim}	
\end{algorithm}
%\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The alogrithm will be similar to the PSO algorithm  in section \ref{sec:ParticleSwarmAlgorithm}.

@@

%Step & Discription of Step & Input & Steps & Output \\ \hline
Stroke 1 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} & 
\begin{scriptsize}
Stroke has 62 points = [P(198.0,153.0), P(197.0,150.0), P(197.0,145.0), P(198.0,140.0), P(203.0,125.0), P(205.0,120.0) , P(205.0,117.0), P(210.0,111.0) , P(214.0,103.0) , P(219.0,97.0), P(225.0,89.0) , P(229.0,85.0) , P(242.0,77.0) , P(247.0,73.0) , P(275.0,64.0) , P(287.0,62.0) , P(325.0,62.0) , P(333.0,63.0) , P(362.0,68.0) , P(372.0,72.0) , P(391.0,77.0) , P(406.0,83.0) , P(424.0,91.0) , P(440.0,100.0), P(474.0,133.0) , P(485.0,149.0) , P(492.0,164.0) , P(499.0,182.0) , P(501.0,191.0) , P(501.0,207.0) , P(499.0,222.0) , P(496.0,239.0) , P(471.0,292.0) , P(467.0,302.0) , P(456.0,315.0) , P(441.0,329.0) , P(427.0,339.0) , P(418.0,344.0) , P(398.0,354.0) , P(336.0,371.0) , P(283.0,376.0) , P(263.0,378.0) , P(242.0,378.0) , P(232.0,376.0) , P(213.0,373.0) , P(194.0,367.0) , P(177.0,362.0) , P(138.0,343.0) , P(129.0,336.0) , P(124.0,331.0) , P(120.0,321.0) , P(115.0,310.0) , P(113.0,297.0) , P(112.0,284.0) , P(114.0,256.0) , P(121.0,219.0) , P(128.0,202.0) , P(137.0,184.0) , P(143.0,172.0) , P(154.0,161.0) , P(172.0,146.0) , P(172.0,146.0)]

@@

	\subfigure[AlgS2 Output]{	\includegraphics{afterDefense/Stroke2Seg2labeled.jpg.eps}	\label{fig:Stroke2Seg2labeled}
	}
	\subfigure[Final Segmentation of Symbol]{		\label{fig:FinalSegmentation}	\includegraphics[scale=0.65]{afterDefense/FinalSegmentation.jpg.eps}
	}
	\caption{Segmentation Ouputs For Second Stroke}
\end{figure}

@@

\begin{figure}
	\centering
		\includegraphics[scale=0.85]{afterDefense/FinalSegmentation.jpg.eps}
	\caption{Best Segmantation }
	\label{fig:segmentatiofinal.jpg}
\end{figure}

@@

	The input of the step is the list of segmentation $L_s$ that is used to compute the feature vector. Figure \ref{fig:segmentatiofinal.jpg} shows the best segmentation as given by the previous step. This step can be summarized as firstly using $L_s$ to computer the feature vector and then introduce this vector the SVM classifier to classify it to one of the trained categories. The details of each step of feature extraction are detailed in the next algorithm. Feature extraction step is divided into four steps where each feature set is computed separately then all sets are appended to construct a final feature vector.

\begin{enumerate}
	\item Computing Structural and geometrical Features (FS1). This step computes number of parallel, perpendicular line and similar structural features. Algorithm \ref{FeatureAlg} is used to determine these features.
	\item Rubine Feature Set (FS2).  Compute Rubine features as in \cite{gestureexample12} 
	\item Statistical Features (FS3) compute the moments as in \cite{zernike61}
	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features.

	\item Introduce the computed feature vector to the classifier and check the label of the detected symbol. 
\end{enumerate}

@@

 FS1 & Structural and geometrical Features & Segment List $L_s$  &  & 
 The final Feature vector of the FS1==>
ature Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468

>>>>

%As you can see form the experiments PSO have proved a spuriously than other system. 
%\section{Future Work}
The system still has some drawbacks; it needs to be improved to handle over traced strokes. Another improvement required is handling of multiple symbols since we only applied the system to single symbols. In addition, full free hand sketch test has not been applied on the system. %sketched

Hence, the next step in this research is to complete the clustering algorithm for fully automated sketch recognition. This will enable testing the system for more than single symbol at a time. The segmentation algorithm may be more powerful if it can segment the stroke into more types of primitives other than the line and curve categorization used in this work. Other area of modification can be the features extraction and classifier. Adding few more spatial and structure features should improve classifications. After segmenting the stroke the system can generate a structural graph from segments and there spatial relations. This graph will be matched with the set of known symbol template to make the identification. Those results along with the SVM classifier results are expected to enhance the system as a whole and achieve even better recognition rates.  %Each new stroke is checked if it can be a symbol or is a part of already drawn un completed symbol.

%drawback of the system is that it takes more time to segment strokes in comparison with other similar algorithms.

\appendix
\chapter{Appendix: Step by Step Example Details}
%\addcontentsline{toc}{chapter}{Appendix: Step by Step Example Details}
\chaptermark{Appendix}
\markboth{Appendix}{Appendix}
\label{ChapterstepExample}%\chapter{Step by Step Example}
%In this chapter a step by step example is presented to clarify the presented system. Let us make an assumption that the user will draw a clock using two strokes. Figure \ref{fig:clock} show the two strokes and how the user draws them. The next steps demonstrate how output of the system after each step. The first section \ref{sec:stepseg} describe how the segmentation is done step by step. Section \ref{sec:steprec} shows the details of feature calculation and the classification process.

@@

%\item Extracting $P_dp$ as in Section \ref{sec:Preprocessing} and Algorithm \ref{extractpdp}
\begin{algorithm}
\caption{Extracting $P_dp$}
\label{extractpdp}
\begin{verbatim}
  1. For each point in Stroke do the following:
  		1.a Compute distance from previous point.
  		1.b Compute Time difference from previous point and add it to Time_list.  
  		1.c Compute Velocity and add it to Vel_list. 
  		1.d Compute Direction and add it to Dir_list.
  	 	1.e Compute Curvature and add it to Cur_list.(Eq.4.1) 
  	end (for loop)
  2. For each  list computed in previous step do the follwing:   
  	 2.a Compute average of all points. 
  	 2.b set average as threshold.  
  	 2.c Create region from continuous points larger than threshold. 
  	 2.d for each region do the following. 
  	 	   2.d.1 find the maximum 
  	 	   2.d.1 Add point to P_dp 
  	  end
     end 
\end{verbatim}	
\end{algorithm}
%\item Ellipse Fitting: which is an attempt to fit as an ellipse as in Section \ref{sec:EllipseDetection} 
%he axes of the ellipse are estimated as the $width/2$ and $height/2$ of the stroke bounding box where $width$ is width and $height$ is height of the bonding box.
\begin{algorithm}
\caption{Ellipse Fitting Algorithm}
\label{ellipseFitAlg}
\begin{verbatim}
1. Initialize the axes of ellipse and center by bounding box information
2. Repeat the following for few loops: 
	2.a Compute error in Equation  4.2 
	2.b Update center of ellipse and axis from error. 
3. Compute percent in Equation 4.4 
4. Computer eff in Equation 4.3 
5. if eff < 0.2  then fit as ellipse otherwise move to segment. 
\end{verbatim}	
\end{algorithm}
%\item Segmentation ALS1 as explained in Section \ref{sec:SwarmSegmentation}. The algorithm will be similar to the PSO algorithm in section \ref{sec:ParticleSwarmAlgorithm}.

@@

%Step & Description of Step & Input & Steps & Output \\ \hline
Stroke 1 & Extracting point and computing boundary data & -- &  see Section \ref{sec:Preprocessing} & 
\begin{scriptsize}
Stroke has 62 points = [P(198.0,153.0), P(197.0,150.0), P(197.0,145.0), P(198.0,140.0), P(203.0,125.0), P(205.0,120.0) , P(205.0,117.0), P(210.0,111.0) , P(214.0,103.0) , P(219.0,97.0), P(225.0,89.0) , P(229.0,85.0) , P(242.0,77.0) , P(247.0,73.0) , P(275.0,64.0) , P(287.0,62.0) , P(325.0,62.0) , P(333.0,63.0) , P(362.0,68.0) , P(372.0,72.0) , P(391.0,77.0) , P(406.0,83.0) , P(424.0,91.0) , P(440.0,100.0), P(474.0,133.0) , P(485.0,149.0) , P(492.0,164.0) , P(499.0,182.0) , P(501.0,191.0) , P(501.0,207.0) , P(499.0,222.0) , P(496.0,239.0) , P(471.0,292.0) , P(467.0,302.0) , P(456.0,315.0) , P(441.0,329.0) , P(427.0,339.0) , P(418.0,344.0) , P(398.0,354.0) , P(336.0,371.0) , P(283.0,376.0) , P(263.0,378.0) , P(242.0,378.0) , P(232.0,376.0) , P(213.0,373.0) , P(194.0,367.0) , P(177.0,362.0) , P(138.0,343.0) , P(129.0,336.0) , P(124.0,331.0) , P(120.0,321.0) , P(115.0,310.0) , P(113.0,297.0) , P(112.0,284.0) , P(114.0,256.0) , P(121.0,219.0) , P(128.0,202.0) , P(137.0,184.0) , P(143.0,172.0) , P(154.0,161.0) , P(172.0,146.0) , P(172.0,146.0)]

@@

	\subfigure[AlgS2 Output]{	\includegraphics{afterDefense/Stroke2Seg2labeled.jpg.eps}	\label{fig:Stroke2Seg2labeled}
	}
	\subfigure[Final Segmentation of Symbol]{		\label{fig:FinalSegmentation}	\includegraphics[scale=0.65]{afterDefense/FinalSegmentation.jpg.eps}
	}
	\caption{Segmentation Outputs For Second Stroke}
\end{figure}

@@

\begin{figure}
	\centering
		\includegraphics[scale=0.85]{afterDefense/FinalSegmentation.jpg.eps}
	\caption{Best Segmentation }
	\label{fig:segmentatiofinal.jpg}
\end{figure}

@@

	The input of the step is the list of segmentation $L_s$ that is used to compute the feature vector. Figure \ref{fig:segmentatiofinal.jpg} shows the best segmentation as given by the previous step. This step can be summarized as firstly using $L_s$ to computer the feature vector and then introduce this vector the SVM classifier to classify it to one of the trained categories. The details of each step of feature extraction are detailed in the next algorithm. Feature extraction step is divided into four steps where each feature set is computed separately then all sets are appended to construct a final feature vector. 
	\begin{enumerate}
	\item Computing Structural and geometrical Features(FS1). This step computes number of parallel, perpendicular line and similar structural features [13 features]. Algorithm \ref{FeatureAlg} in Appendix \ref{ChapterstepExample}  is used to determine these features.
 	\item Rubine Feature Set (FS2) [12 features] .  Compute Rubine features as in \cite{gestureexample12} 
	\item Statistical Features (FS3) compute the moments as in \cite{zernike61} for N=10 there are 32 features.
	\item Global shape properties set (FS4). Firstly compute the convex hull of the stroke points then compute the rest of the features [13 features]. 
	\item Introduce the computed feature vector (All 70 features) to the classifier and check the label of the detected symbol. 
\end{enumerate}

@@

 FS1 & Structural and geometrical Features & Segment List $L_s$  &  & 
 The final Feature vector of the FS1==>
  Number of primitives 3.0   ,    Feature Number of segments 9.0   ,    Feature Line count 6.0   ,    Feature Curves Count 3.0   ,    Feature Ellipse Count 1.0   ,    Feature Intersection T type -1.0   ,    Feature Intersection L type 4.0   ,    Feature Intersection X type -1.0   ,    Feature Parallel Count 1.0   ,    Feature PerPendicular Count -1.0   ,    Feature Intersection Lines Count 4.0   ,    Feature Min Radius 1.7976931348623157E308   ,    Feature Max Radius 231.70851748053468

### 2009-10-13 trunk/draft/Results.tex
FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL FINAL

Several experiments are presented to test the different segmentation algorithms on the three data sets used in our system. Algorithms (\textsl{AlgS1}, \textsl{AlgS2} and \textsl{Alg3} (in section \ref{sec:BenchMarckAlgorithm})) were tested individually with and without the ellipse fitting module. The final system (Choosing the Minimum error of (\textsl{AlgS1}, \textsl{AlgS2}) was evaluated on the three dataset(HS, EL and LD datasets ). Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} shows the accuracy achieved by each algorithm. Theses experiments were done using the whole dataset as 75\% of the data as test set and 25\% as training. The results shows that \textsl{AlgS2} with the ellipse fitting module gives better recognition rate across all datasets (Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} ). Even though, \textsl{AlgS1} gives better result than \textsl{Alg3} in both EL-DB and HS-DB it performs poorly in LD-DB due to the high number of curves in every symbol in this dataset. These result are obtained using different number of samples per category as shown in Figures \ref{fig:LDRes}, \ref{fig:ElRes}. 
% i need to organize this as following. 
% number 1. differnt algorithm on different data 
% which algortihm is better oon which data 
% on ld and el  the alg2  is better,

@@

 The main features affected by the segmentation stage are the geometric features computed in\textbf{FS1}. Hence those are the features we used to estimate the efficiency of the segmentation algorithm. The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly} . The figure shows that \textsl{AlgS2} achieve better results.  %effect of the segmentation algorithm 
 %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block. The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}). Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features. 
 \begin{figure}
	\centering
		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
	\caption{The effect of segmentation on Hs-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeatonly}
\end{figure}  
% \begin{figure}
%	\centering
%		\includegraphics[scale=0.3]{paperImages/featureFS1only.jpg.eps}
%	\caption{The effect of segmentation on LD-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
%	\label{fig:LDtestFeatonly}
%\end{figure}  
 \begin{figure}
	\centering
		\includegraphics[scale=0.8]{results/F1EL.eps}
	\caption{The effect of segmentation on EL-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:ELtestFeatonly}
\end{figure}  
%
% \begin{figure}
%	\centering
%		\includegraphics[scale=0.5]{results/testFeat.eps}
%	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{Hs-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
%	\label{fig:testFeaturesAllHS}
%\end{figure}  
%
% \begin{figure}
%	\centering
%		\includegraphics[scale=0.5]{results/testF1.eps}
%	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{MD-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
%	\label{fig:testFeaturesAllMD}
%\end{figure}  
%Another expirements where conducted to evaluates the worth of an attribute. The worth of each attribute is measured using two different attribure measures. The first is by measuring the gain ratio with respect to the class, where the gain ration is computed as $Gain_{ration}(class,attribute)=\frac{H(class)-H(class|Attribure}{H(Attribute)}$. The second method is using

>>>>

Several experiments are presented to test the different segmentation algorithms on the three data sets used in our system. Algorithms (\textsl{AlgS1}, \textsl{AlgS2} and \textsl{Alg3} (in section \ref{sec:BenchMarckAlgorithm})) were tested individually with and without the ellipse fitting module. The final system (Choosing the Minimum error of (\textsl{AlgS1}, \textsl{AlgS2}) was evaluated on the three dataset (HS, EL and LD datasets). Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} shows the accuracy achieved by each algorithm. These experiments were done using the whole dataset as 75\% of the data as test set and 25\% as training. The results shows that \textsl{AlgS2} with the ellipse fitting module gives better recognition rate across all datasets (Figure \ref{fig:test1},\ref{fig:testEL} and \ref{fig:testLD} ). Even though, \textsl{AlgS1} gives better result than \textsl{Alg3} in both EL-DB and HS-DB it performs poorly in LD-DB due to the high number of curves in every symbol in this dataset. These result are obtained using different number of samples per category as shown in Figures \ref{fig:LDRes}, \ref{fig:ElRes}. 
% i need to organize this as following. 
% number 1. differnt algorithm on different data 
% which algortihm is better oon which data 
% on ld and el  the alg2  is better,

@@

 The main features affected by the segmentation stage are the geometric features computed in\textbf{FS1}. Hence those are the features we used to estimate the efficiency of the segmentation algorithm. The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly} . The figure shows that \textsl{AlgS2} achieve better results.  %effect of the segmentation algorithm 
 %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block. The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}). Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features. 
 \begin{figure}
	\centering
		\includegraphics[scale=0.4]{paperImages/featureFS1only.jpg.eps}
	\caption{The effect of segmentation on Hs-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeatonly}
\end{figure}  
% \begin{figure}
%	\centering
%		\includegraphics[scale=0.3]{paperImages/featureFS1only.jpg.eps}
%	\caption{The effect of segmentation on LD-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
%	\label{fig:LDtestFeatonly}
%\end{figure}  
 \begin{figure}
	\centering
		\includegraphics[scale=0.8]{results/F1EL.eps}
	\caption{The effect of segmentation on EL-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:ELtestFeatonly}
\end{figure}  
%
% \begin{figure}
%	\centering
%		\includegraphics[scale=0.5]{results/testFeat.eps}
%	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{Hs-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
%	\label{fig:testFeaturesAllHS}
%\end{figure}  
%
% \begin{figure}
%	\centering
%		\includegraphics[scale=0.5]{results/testF1.eps}
%	\caption{Feature Comparison} The effect of different features on accuracy on \textsl{MD-DB}.  %The graph shows the recognition rate of each symbol using different algorithms. 
%	\label{fig:testFeaturesAllMD}
%\end{figure}  
%Another experiments where conducted to evaluates the worth of an attribute. The worth of each attribute is measured using two different attribute measures. The first is by measuring the gain ratio with respect to the class, where the gain ration is computed as $Gain_{ration}(class,attribute)=\frac{H(class)-H(class|Attribure}{H(Attribute)}$. The second method is using
